---
title: "Data cleaning"
author: "Sitian Zhou, Pei Tian"
date: "`r Sys.Date()`"
output: html_document
---
```{r setup, include=F}
knitr::opts_chunk$set(message = F, warning = F)
```

**Delivering dataset preprocessing to ensure the feasibility of merging different datasets for downstream analysis.**

```{r lib-import}
library(tidyverse)
library(rvest)
library(httr)
library(readxl)
library(janitor)
library(fuzzyjoin)
```

## Country code

The country code dataset contains country names along with their corresponding standard two-letter, three-letter, and numeric codes. The dataset can be accessed [here](https://www.kaggle.com/datasets/juanumusic/countries-iso-codes). 

**Key steps in cleaning this dataset**

* Renaming some countries
* Variable selection

```{r country-code}
country_code <-
  read_csv("../data/raw/country_code.csv") |> 
  clean_names() |> 
  rename(
    "country_name" = "english_short_name_lower_case"
  ) |> 
  mutate(
    country_name = recode(country_name, 
                          "United States Of America" = "United States",
                          "Virgin Islands, British" = "British Virgin Islands",
                          "Korea, Republic of (South Korea)" = "South Korea",
                          "Virgin Islands, U.S." = "U.S. Virgin Islands",
                          "Tanzania, United Republic of" = "Tanzania",
                          "Turks and Caicos" = "Turks and Caicos Islands",
                          "Macao" = "Macau")) |> 
  select(country_name, alpha_3_code)

```


### Billionaires 2010-2023

The raw dataset contains information on global billionaires from 1997 to 2023. It offers a glimpse into the distribution of wealth, industries of operation, and demographic profiles of billionaires on a global scale. The dataset is available [here](https://www.kaggle.com/datasets/guillemservera/forbes-billionaires-1997-2023?select=all_billionaires_1997_2023.csv).

**Key steps in cleaning this dataset**

* Filtering for data from 2010 to 2023
* Converting `net_worth` to numerical value
* Variable selection
* Merging with `country_code` dataset for further analysis

```{r recent-ten-years-billionaires}
bil_2010_2023 <-
  read_csv("../data/raw/billionaires_1997_2023.csv") |> 
  filter(year >= 2010)

bil_2010_2023_clean <-
  bil_2010_2023 |> 
  mutate(
    net_worth = str_replace_all(net_worth, " B", ""),
    net_worth = as.numeric(net_worth), 
    industries = str_replace_all(business_industries, "[\\['\\]]", ""),
    country_of_residence = 
      recode(country_of_residence, 
             "Eswatini (Swaziland)" = "Swaziland",
             "Scotland" = "United Kingdom",
             "Czechia" = "Czech Republic",
             "Hong Kong SAR" = "Hong Kong"),
    industries = 
      recode(industries,
             "Fashion and Retail" = "Fashion & Retail",
             "Finance and Investments" = "Finance & Investments",
             "Food and Beverage" = "Food & Beverage",
             "Healthcare" = "Health care",
             "Media" = "Media & Entertainment")) |> 
  left_join(country_code, c("country_of_residence" = "country_name")) |> 
  select(-c(month, rank, last_name, first_name, birth_date, business_category,
            business_industries, organization_name, position_in_organization))

```

### Billionaires 2023 with GDP data

The billionaires 2023 dataset includes statistics on global billionaires, such as information about their wealth, industries, and personal details. This dataset also contains more detailed country information of which each billionaire resides, which is useful for the following analysis. The dataset is available [here](https://www.kaggle.com/datasets/nelgiriyewithana/billionaires-statistics-dataset).

**Key steps in cleaning this dataset**

* Cleaning and changing the unit of `final_worth` to billion
* Cleaning and changing the unit of `gdp_country` to trillion
* Recoding `gender` variable
* Variable selection

```{r}
bil_gdp_2023 <-
  read_csv("../data/raw/billionaires_2023.csv")

bil_gdp_2023_clean <-
  bil_gdp_2023 |> 
  clean_names() |> 
  mutate(
    net_worth = final_worth / 1000,
    gdp_country = str_replace_all(gdp_country, "[$,]", ""),
    gdp_country = as.numeric(gdp_country) / 1e12,
    gender = 
      case_match(
        gender,
        "F" ~ "Female",
        "M" ~ "Male")) |>
  select(net_worth, full_name = person_name, age, gender, 
         country_of_citizenship, country_of_residence = country, 
         city_of_residence = city, industries, self_made, cpi_country, 
         cpi_change_country, gdp_country, life_expectancy_country)

```


## Region-level GDP 

### Basic cleaning

The GDP dataset encompasses GDP information spanning 262 distinct countries or regions from 1960 to 2022. The dataset can be downloaded [here](https://www.kaggle.com/datasets/willianoliveiragibin/world-gdp-by-country-region-and-income-group?select=gdp_data.csv).

**Key steps in cleaning this dataset**

* Filtering for data starts from 2010
* Changing the unit of GDP to trillion
* Variable renaming and selection

```{r country-gdp-data}
country_gdp <-
  read_csv("../data/raw/country_gdp.csv")

country_gdp_clean <-
  country_gdp |> 
  filter(year >= 2010) |> 
  mutate(gdp = value / 1e12,
         name = country_name, 
         code = country_code) |> 
  select(name, code, year, gdp)
  
```

### Supplementary dataset

Given the issue that the region `Taiwan` is missing GDP data in raw dataset, we used a supplementary dataset to remedy it.

#### Taiwan

The Taiwan GDP data can be accessed [here](https://countryeconomy.com/gdp/taiwan). This table contains Taiwan GDP data from 1960 to 2022.

**Key steps in cleaning this dataset**

* Scapping GDP data from specific website
* Extracting and filtering GDP data
```{r tw-gdp-data}
fetch_tw_gdp = function(){
  url = "https://countryeconomy.com/gdp/taiwan"
  tw_gdp_html = read_html(url)
  mydata <- tw_gdp_html|> html_table()
  return(mydata[[1]])
}

extract_gdp = function(string){
  str_vec = str_extract(string, "\\d*,\\d*") |> 
    str_split(",") |> 
    nth(1)
  s = ""
  for(e in str_vec){
    s = str_c(s, e)
  }
  return(s)
}

tw_gdp = fetch_tw_gdp() |>
  janitor::clean_names() |>
  mutate(gdp = map(annual_gdp_2, extract_gdp) |> as.numeric(),
         year = date |> as.numeric(), 
         name = "Taiwan, China",
         code = "TWN",
         gdp = gdp/1e6) |>
  filter(year >= 2010) |>
  select(name, code, year, gdp)
```

### Merge GDP data
* Combining Taiwan GDP data with the main GDP dataset
```{r merge-region-gdp}
country_gdp_clean = bind_rows(country_gdp_clean, tw_gdp)
```


## Industry-level GDP for US

### Basic cleaning

This dataset contains GDP data for each industry in US from 2017 to 2022. The dataset can be downloaded [here](https://www.bea.gov/data/gdp/gdp-industry).

**Key steps in cleaning this dataset**

* Renaming variable names
* Tidying the data

```{r import-industry-data}
indus_gdp <-
  read_excel("../data/raw/usa_industry_gdp.xlsx", sheet = 18, skip = 4)

indus_gdp_clean <-
  indus_gdp |> 
  slice(3:30) |> 
  rename(
    "industries" = "...2",
    "2017" = "...3",
    "2018" = "...4",
    "2019" = "...5",
    "2020" = "...6",
    "2021" = "...7",
    "2022" = "2020...8") |>
  filter(!(industries %in% c("Finance, insurance, real estate, rental, and leasing",
                            "Educational services, health care, and social assistance",
                      "Arts, entertainment, recreation, accommodation, and food services"))) |> 
  select(industries, `2017`:`2022`) |> 
  pivot_longer(`2017`:`2022`, names_to = "year", values_to = "industry_gdp") |> 
  mutate(year = as.numeric(year),
         industry_gdp = industry_gdp / 1000)

```

### Merge industry-level GDP & billionaire

* Creating two summary tables with industry names
* Extracting keywords for each industry
* Using regex to perform inexact matches for industries

```{r reg-match-industry}
# extract key words to merge two datasets by industries variable
#
# industry names from industry_gdp_clean
df1 <- indus_gdp_clean |> select(industries) |> unique()
# industry names from bil_2013_2023_clean (only for USA from 2017 to 2022!)
df2 <- 
  bil_2010_2023_clean |> 
  filter(country_of_residence == "United States" & year >= 2017 & year <= 2022) |> 
  select(industries) |> 
  drop_na() |> 
  unique() |> 
  mutate(
    categories = industries,
    # rename some industries for better match results
    categories = recode(categories, 
                        "Technology" = "Information",
                        "Logistics" = "Transportation and warehousing")) |> 
  # extract keywords for each industry
  separate(categories, into = c("word1", "word2"), sep = " & ") |> 
  pivot_longer(
    word1:word2,
    names_to = "order",
    values_to = "keywords"
  ) |> 
  drop_na(keywords)

# use regex to perform inexact matching
reg_match <-
  regex_inner_join(df1, df2, by=c("industries" = "keywords"), ignore_case = TRUE) |> 
  distinct(industries.y, .keep_all = TRUE) |>
  # removed service industry bc it's too general
  filter(industries.y != "Service") |> 
  select(industries.x, industries.y)
```

### Merge GDP & billionaire

* Joining the dataset for billionaires from 2010 to 2023 with the country GDP dataset
* Variable renaming and selection

```{r merge-2}
bil_gdp_2010_2023 <-
  left_join(bil_2010_2023_clean, country_gdp_clean, 
            by = join_by(alpha_3_code == code, year == year), multiple = "all") |>
  mutate(region_gdp = gdp, region_code = alpha_3_code) |>
  select(-c(gdp, alpha_3_code, name))
```

## USA industry dataset

* Filtering the billionaires dataset for only billionaires reside in US from year 2017 to 2022
* Merging the filtered dataset with industry-level GDP dataset
* Variable selection

```{r prepare-USA-industry-dataset}
bil_gdp_indus_usa <-
  bil_gdp_2010_2023 |> 
  filter(country_of_residence == "United States" & year >= 2017 & year <= 2022) |> 
  left_join(reg_match, c("industries" = "industries.y")) |> 
  left_join(indus_gdp_clean, c("industries.x" = "industries", "year" = "year")) |> 
  select(-industries.x, -region_code)

```


* Save the cleaned datasets

```{r save-data}
### save useful files & add some description
write_csv(bil_gdp_2010_2023, "../data/tidy/billionaire_gdp.csv")
write_csv(bil_gdp_indus_usa, "../data/tidy/billionaire_gdp_indus_usa.csv")
write_csv(country_gdp_clean, "../data/tidy/gdp.csv")
write_csv(bil_gdp_2023_clean, "../data/tidy/billionaire_2023.csv")
```

## Description for tidy data
Describe dataset saved for downstream analysis.


**billionaire_gdp.csv**

This dataset covers global billionaires from 2010 to 2023, detailing their wealth, personal information, and country-specific GDP.

`year`: The year in which the data was collected

`net_worth`: The net worth of the billionaire (in billions of dollars)

`full_name`, `age`, and `gender`: The name, age, and gender of the billionaire, respectively

`country_of_citizenship`, `country_of_residence`, and `city_of_residence`: The residence details spanning country, city, and citizenship of the billionaire

`self_made`: Whether or not the billionaire's wealth is self-made or inherited

`wealth_status`: Current status of the billionaire's wealth, whether it's rising, stable, or declining

`industries`: The industries associated with the billionaire's business interests

`region_gdp`: The real GDP (in trillions of dollars) for each country or region the billionaire resides

`region_code`: The three-letter code of the corresponding country or region the billionaire resides


**billionaire_2023.csv**

This dataset contains information for global billionaires in 2023, including their wealth, personal details, and country-specific economic status.

`year`: The year in which the data was collected

`net_worth`: The net worth of the billionaire (in billions dollars)

`full_name`, `age`, and `gender`: The name, age, and gender of the billionaire, respectively

`country_of_citizenship`, `country_of_residence`, and `city_of_residence`: The residence details spanning country, city, and citizenship of the billionaire

`self_made`: Whether or not the billionaire is self-made

`cpi_country` and `cpi_change_country`: Consumer Price Index (CPI) and its change for the billionaire's country

`gdp_country`: The real GDP (in trillions dollars) for each country the billionaire resides

`life_expectancy_country`: Life expectancy in the country the billionaire reside


**billionaire_gdp_indus_usa.csv**

This dataset combines information on billionaires residing in the United States from 2017 to 2022, with the annual GDP and the GDP for various industries in the United States.

`year`: The year in which the data was collected

`net_worth`: The net worth of the billionaire (in billions dollars)

`full_name`, `age`, and `gender`: The name, age, and gender of the billionaire, respectively

`country_of_citizenship`, `country_of_residence`, and `city_of_residence`: The residence details spanning country, city, and citizenship of the billionaire

`self_made`: Whether or not the billionaire's wealth is self-made or inherited

`wealth_status`: Current status of the billionaire's wealth

`industries`: The industries associated with the billionaire's business interests

`region_gdp`: The real GDP (in trillions dollars) for each country or region the billionaire resides

`region_code`: The three-letter code of the corresponding country or region the billionaire resides

`industry_gdp`: The annual GDP (in trillions dollars) for the industry

**gdp.csv**

This dataset includes information of GDP data of each region.

`name`: region name (country or SAR(special administration area))

`code`: region code

`year`: year

`gdp`: GDP value (trillion as unit)

