---
title: "data_clean"
author: "Sitian Zhou"
date: "2023-11-07"
output: html_document
---

```{r}
library(tidyverse)
library(readxl)
library(janitor)
```

## country code

```{r}
country_code <-
  read_csv("../data/country_code.csv") |> 
  clean_names() |> 
  rename(
    "country_name" = "english_short_name_lower_case"
  ) |> 
  mutate(
    country_name = recode(country_name, 
                          "United States Of America" = "United States",
                          "Virgin Islands, British" = "British Virgin Islands",
                          "Korea, Republic of (South Korea)" = "South Korea",
                          "Virgin Islands, U.S." = "U.S. Virgin Islands",
                          "Tanzania, United Republic of" = "Tanzania",
                          "Turks and Caicos" = "Turks and Caicos Islands",
                          "Macao" = "Macau")) |> 
  select(country_name, alpha_3_code)


```


## Billionaires 2013-2023

* filtered for data from 2013 to 2023
* converted `net_worth` to numerical value
* selected/renamed variables

```{r}
bil_2013_2023 <-
  read_csv("../data/billionaires_1997_2023.csv") |> 
  filter(year >= 2013)

bil_2013_2023_clean <-
  bil_2013_2023 |> 
  mutate(
    net_worth = str_replace_all(net_worth, " B", ""),
    net_worth = as.numeric(net_worth), 
    industries = str_replace_all(business_industries, "[\\['\\]]", ""),
    country_of_residence = 
      recode(country_of_residence, 
             "Eswatini (Swaziland)" = "Swaziland",
             "Scotland" = "United Kingdom",
             "Czechia" = "Czech Republic")) |> 
  left_join(country_code, c("country_of_residence" = "country_name")) |> 
  select(-c(month, rank, last_name, first_name, birth_date, business_category,
            business_industries, organization_name, position_in_organization))
  
```


## Billionaires 2023 only

* changed the unit of `final_worth` to billion; result column as `net_worth`
* cleaned `gdp_country`; changed the unit to trillion
* recoded `gender`
* selected/renamed variables

```{r}
bil_gdp_2023 <-
  read_csv("../data/billionaires_2023.csv")

bil_gdp_2023_clean <-
  bil_gdp_2023 |> 
  clean_names() |> 
  mutate(
    net_worth = final_worth / 1000,
    gdp_country = str_replace_all(gdp_country, "[$,]", ""),
    gdp_country = as.numeric(gdp_country) / 1e12,
    gender = 
      case_match(
        gender,
        "F" ~ "Female",
        "M" ~ "Male")) |>
  select(net_worth, full_name = person_name, age, gender, 
         country_of_citizenship, country_of_residence = country, 
         city_of_residence = city, industries, cpi_country, 
         cpi_change_country, gdp_country, life_expectancy_country, 
         latitude_country, longitude_country)

```

## country GDP dataset

* filtered for data starts from 2013
* changed the unit of gdp to trillion

```{r}
country_gdp <-
  read_csv("../data/country_gdp.csv")

country_gdp_clean <-
  country_gdp |> 
  filter(year >= 2013) |> 
  mutate(gdp = value / 1e12) |> 
  select(year, everything(), -value)
  
```



## joined dataset for billionaires 2013-2023 w/ country GDP info

some problems:
1. Hong Kong, Taiwan, Macau: missing GDP data


```{r}
bil_gdp_2013_2023 <-
  left_join(bil_2013_2023_clean, country_gdp_clean, 
            c("alpha_3_code" = "country_code", "year" = "year")) |> 
  select(-c(alpha_3_code, country_name))
  
```


## industry GDP for US

(not started)

```{r}
industry_gdp <-
  read_excel("../data/usa_industry_gdp.xlsx", sheet = 18, skip = 4)
```


