---
title: "data_clean"
author: "Sitian Zhou"
date: "2023-11-07"
output: html_document
---

```{r}
library(tidyverse)
library(readxl)
library(janitor)
library(fuzzyjoin)
```

## country code


Changed some country names so they match the country name in the 

```{r}
country_code <-
  read_csv("../data/country_code.csv") |> 
  clean_names() |> 
  rename(
    "country_name" = "english_short_name_lower_case"
  ) |> 
  mutate(
    country_name = recode(country_name, 
                          "United States Of America" = "United States",
                          "Virgin Islands, British" = "British Virgin Islands",
                          "Korea, Republic of (South Korea)" = "South Korea",
                          "Virgin Islands, U.S." = "U.S. Virgin Islands",
                          "Tanzania, United Republic of" = "Tanzania",
                          "Turks and Caicos" = "Turks and Caicos Islands",
                          "Macao" = "Macau")) |> 
  select(country_name, alpha_3_code)


```


## Billionaires 2013-2023

* filtered for data from 2013 to 2023
* converted `net_worth` to numerical value
* selected/renamed variables
* merged with `country_code` dataset for further analysis

```{r}
bil_2013_2023 <-
  read_csv("../data/billionaires_1997_2023.csv") |> 
  filter(year >= 2013)

bil_2013_2023_clean <-
  bil_2013_2023 |> 
  mutate(
    net_worth = str_replace_all(net_worth, " B", ""),
    net_worth = as.numeric(net_worth), 
    industries = str_replace_all(business_industries, "[\\['\\]]", ""),
    country_of_residence = 
      recode(country_of_residence, 
             "Eswatini (Swaziland)" = "Swaziland",
             "Scotland" = "United Kingdom",
             "Czechia" = "Czech Republic"),
    industries = 
      recode(industries,
             "Fashion and Retail" = "Fashion & Retail",
             "Finance and Investments" = "Finance & Investments",
             "Food and Beverage" = "Food & Beverage",
             "Healthcare" = "Health care",
             "Media" = "Media & Entertainment")) |> 
  left_join(country_code, c("country_of_residence" = "country_name")) |> 
  select(-c(month, rank, last_name, first_name, birth_date, business_category,
            business_industries, organization_name, position_in_organization))

```


## Billionaires 2023 only

* changed the unit of `final_worth` to billion; result column as `net_worth`
* cleaned `gdp_country`; changed the unit to trillion
* recoded `gender`
* selected/renamed variables

```{r}
bil_gdp_2023 <-
  read_csv("../data/billionaires_2023.csv")

bil_gdp_2023_clean <-
  bil_gdp_2023 |> 
  clean_names() |> 
  mutate(
    net_worth = final_worth / 1000,
    gdp_country = str_replace_all(gdp_country, "[$,]", ""),
    gdp_country = as.numeric(gdp_country) / 1e12,
    gender = 
      case_match(
        gender,
        "F" ~ "Female",
        "M" ~ "Male")) |>
  select(net_worth, full_name = person_name, age, gender, 
         country_of_citizenship, country_of_residence = country, 
         city_of_residence = city, industries, cpi_country, 
         cpi_change_country, gdp_country, life_expectancy_country, 
         latitude_country, longitude_country)

```

## country GDP dataset

* filtered for data starts from 2013
* changed the unit of gdp to trillion

```{r}
country_gdp <-
  read_csv("../data/country_gdp.csv")

country_gdp_clean <-
  country_gdp |> 
  filter(year >= 2013) |> 
  mutate(gdp = value / 1e12) |> 
  select(year, everything(), -value)
  
```



## joined dataset for billionaires 2013-2023 w/ country GDP info

some problems:
* Hong Kong, Taiwan, Macau: missing GDP data
* missing GDP data for 2023 - GDP in `bil_gdp_2023_clean` inaccurate(?)

```{r}
bil_gdp_2013_2023 <-
  left_join(bil_2013_2023_clean, country_gdp_clean, 
            c("alpha_3_code" = "country_code", "year" = "year")) |> 
  select(-c(alpha_3_code, country_name))
```


## industry GDP for US

* renamed variable names
* tidied the data

```{r}
indus_gdp <-
  read_excel("../data/usa_industry_gdp.xlsx", sheet = 18, skip = 4)

indus_gdp_clean <-
  indus_gdp |> 
  slice(3:30) |> 
  rename(
    "industries" = "...2",
    "2017" = "...3",
    "2018" = "...4",
    "2019" = "...5",
    "2020" = "...6",
    "2021" = "...7",
    "2022" = "2020...8") |>
  filter(!(industries %in% c("Finance, insurance, real estate, rental, and leasing",
                            "Educational services, health care, and social assistance",
                      "Arts, entertainment, recreation, accommodation, and food services"))) |> 
  select(industries, `2017`:`2022`) |> 
  pivot_longer(`2017`:`2022`, names_to = "year", values_to = "industry_gdp") |> 
  mutate(year = as.numeric(year))
```

## Merge datasets by industries

* created two summary tables with industry names
* used regex to match the two

```{r}
# extract key words to merge two datasets by industries variable
#
# industry names from industry_gdp_clean
df1 <- indus_gdp_clean |> select(industries) |> unique()
# industry names from bil_2013_2023_clean
df2 <- bil_2013_2023_clean |> select(industries) |> drop_na() |> 
  unique() |> 
  mutate(
    categories = industries,
    # rename some industries for better match results
    categories = recode(categories, 
                        "Technology" = "Information",
                        "Politics" = "State and local",
                        "Logistics" = "Transportation and warehousing",
                        "Medicine" = "Health care")) |> 
  # extract keywords for each industry
  separate(categories, into = c("word1", "word2"), sep = " & ") |> 
  pivot_longer(
    word1:word2,
    names_to = "order",
    values_to = "keywords"
  ) |> 
  drop_na(keywords)

# use regex to perform inexact matching
reg_match <-
  regex_inner_join(df1, df2, by=c("industries" = "keywords"), ignore_case = TRUE) |> 
  distinct(industries.y, .keep_all = TRUE) |>
  # removed service industry bc it's too general
  filter(industries.y != "Service") |> 
  select(industries.x, industries.y)
```


## final dataset

```{r}
bil_gdp_indus_2013_2023 <-
  bil_gdp_2013_2023 |> 
  left_join(reg_match, c("industries" = "industries.y")) |> 
  left_join(indus_gdp_clean, c("industries.x" = "industries", "year" = "year")) |> 
  select(-industries.x)
```

