---
title: "Geographic Analysis"
author: "Pei Tian"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=F}
knitr::opts_chunk$set(message = F, warning = F)
```

```{r lib-import}
library(tidyverse)
library(leaflet)
library(sf)
library(rnaturalearth)
library(plotly)

theme_set(
  theme_bw() +
  theme(plot.title = element_text(hjust = 0.5),
        text = element_text(family = "Helvetica")
  )
)

```

## Purposes
Interactive visualization for top-ranked billionaires in different regions (countries, continents)

Interactive visualization for GDP of different countries in different years

Compare billionaire distribution in different regions (such as average wealth, gender composition)

```{r data-preprocessing}
bil_gdp = read_csv("../data/tidy/billionaire_gdp.csv") 
bil_gdp = bil_gdp |>
  select(-wealth_status) |>
  mutate(gender = factor(ifelse(is.na(gender), "Unknown", gender)),
         year = as.integer(year),
         age = as.integer(age),
         country_of_residence = factor(country_of_residence),
         country_of_citizenship = factor(country_of_citizenship),
         region_code = factor(region_code),
         industries = factor(ifelse(is.na(industries), "Unknown", industries)),
         self_made = factor(ifelse(is.na(self_made), "Unknown", self_made))
         )

gdp_data = read_csv("../data/raw/gdp_worldbank.csv", skip = 4) |>
  janitor::clean_names() |>
  select(country_name, country_code, starts_with("x")) |>
  pivot_longer(cols = starts_with("x"), names_to = "year", names_prefix = "x", values_to = "value") |>
  mutate(gdp = value/1e12, 
         year = as.integer(year),
         name = country_name, 
         code = country_code) |>
  filter(year >= 2010) |>
  select(name, code, year, gdp)

```

## Interactive Map
Input: `year`
- top1 Billionaire in each country as label
- GDP reveal as color

```{r geo-data}
world_polygon = ne_countries(scale = "medium", returnclass = "sf")
world_point = st_point_on_surface(world_polygon)

```

```{r geo-view}
target_year = 2022
gdp_viz = gdp_data |>
  filter(year == target_year) |>
  left_join(world_polygon, by = join_by(code == adm0_a3)) |>
  st_as_sf()

bil_viz = bil_gdp |>
  filter(year == target_year) |>
  group_by(region_code, year) |>
  arrange(desc(net_worth)) |>  # Arrange in descending order based on 'value'
  mutate(rnk = row_number()) |>  # Add rank labels
  filter(rnk <= 1) |>
  mutate(info = paste(paste0("Top 1 Billionaire: ", full_name), industries, paste0(net_worth, " billion"), sep = ", ")) |>
  left_join(world_point, by = join_by(region_code == adm0_a3)) |>
  st_as_sf()

leaflet() |>
  addProviderTiles("CartoDB.Positron") |>  # You can choose different tile providers
  addPolygons(data = gdp_viz, 
              fillColor = ~colorQuantile("YlOrRd", gdp)(gdp), 
              fillOpacity = 0.3, color = "white", weight = 1,
              label = ~paste("GDP: ", round(gdp, 3), " trillion"),
              highlightOptions = highlightOptions(
                color = "black", weight = 2, bringToFront = TRUE)) |>
  addMarkers(data = bil_viz,
             label = ~paste("Region: ", region_code),
             popup = ~info
  ) |>
  setView(lng = 0, lat = 30, zoom = 2)
```

## Chart
### Global analysis
1. Barplot: count in different region
  Input: year, age
  View: 
  - gender
  - industry
  - self-made
  
```{r inter-data}
target_year = 2022
target_age_range = c(1, 100)
inter_data = bil_gdp |>
  filter(year == target_year & age >= target_age_range[1] & age <= target_age_range[2])

```

```{r count-view}
count_by_region = inter_data |>
  group_by(region_code) |>
  summarise(count = n()) |>
  mutate(region_code = fct_reorder(region_code, count))
count_view = count_by_region |>
  filter(count >= 3 & !is.na(region_code)) |>
  ggplot(aes(x = region_code, y = count, fill = "tomato")) +
  geom_bar(stat = "identity") +
  labs(title = "Billionaire Count", x = "Region", y = "Count") +
  theme(axis.text.x = element_text(angle = -90, hjust = 1), 
        legend.position = "none")
ggplotly(count_view)
```

```{r gender-view}
count_by_gender = inter_data |>
  group_by(region_code, gender) |>
  summarise(gender_count = n(), .groups = "drop_last") |>
  left_join(count_by_region, by = join_by(region_code == region_code)) |>
  filter(count >= 3 & !is.na(region_code) ) |>
  mutate(prct = gender_count / count * 100)
gender_view = count_by_gender |>
  ggplot(aes(x = region_code, y = prct, fill = gender)) +
  geom_bar(stat = "identity") +
  scale_fill_manual(values = c("tomato", "lightblue", "grey")) +
  labs(title = "Gender Distribution", x = "Region", y = "Precent") +
  theme(axis.text.x = element_text(angle = -90, hjust = 1)) 
ggplotly(gender_view)
```

```{r self-made-view}
count_by_self_made = inter_data |>
  group_by(region_code, self_made) |>
  summarise(self_made_count = n(), .groups = "drop_last") |>
  left_join(count_by_region, by = join_by(region_code == region_code)) |>
  filter(count >= 3 & !is.na(region_code) ) |>
  mutate(prct = self_made_count / count * 100)
self_made_view = count_by_self_made |>
  ggplot(aes(x = region_code, y = prct, fill = self_made)) +
  geom_bar(stat = "identity", alpha = .8) +
  scale_fill_manual(values = c("orange", "violet", "grey"), 
                    labels = levels(count_by_self_made$self_made)) +
  labs(title = "Self-made Distribution", x = "Region", y = "Precent") +
  theme(axis.text.x = element_text(angle = -90, hjust = 1))
ggplotly(self_made_view)
```

```{r industries-view}
count_by_industries = inter_data |>
  group_by(region_code, industries) |>
  summarise(industries_count = n(), .groups = "drop_last") |>
  left_join(count_by_region, by = join_by(region_code == region_code)) |>
  filter(count >= 3 & !is.na(region_code) ) |>
  mutate(prct = industries_count / count * 100)
industries_view = count_by_industries |>
  ggplot(aes(x = region_code, y = prct, fill = industries)) +
  geom_bar(stat = "identity", alpha = .8) +
  labs(title = "Industry Distribution", x = "Region", y = "Precent") +
  theme(axis.text.x = element_text(angle = -90, hjust = 1)) +
  viridis::scale_fill_viridis(
    name = "industries", 
    discrete = TRUE,
    option = "H"
  )
ggplotly(industries_view)
```



### Region-specific analysis

2. Specific country: 
  Input: country
  - GDP curve
  - gender, self-made, industry: pie
  - age: histogram
  - wealth: top rank
  
#### Charts
```{r gdp-curve}
code = c("USA", "CHN")
region_select_all = bil_gdp |> 
  filter(region_code %in% code)
region_select = region_select_all |>
  filter(year == target_year)
region_select_all |>
  select(year, region_gdp, region_code) |>
  distinct() |>
  ggplot(aes(x = year, y = region_gdp, color = region_code)) +
  geom_line() +
  geom_point() +
  labs(x = "Year", y = "GDP(Billion)", title = "GDP in recent years") 
```
```{r rnk-view}
rnk_thres = c(1, 10)
rnk_bar = region_select |>
  group_by(region_code) |>
  arrange(desc(net_worth)) |>  # Arrange in descending order based on 'value'
  mutate(rnk = row_number()) |>  # Add rank labels
  filter(rnk <= rnk_thres[2] & rnk >= rnk_thres[1]) |>  # Select top 10 values
  select(region_code, net_worth, full_name, rnk) |>
  ggplot(aes(x = rnk, y = net_worth, fill = region_code, label = full_name)) +
  geom_bar(stat = "identity", position = "dodge", color = "white") +
  scale_x_continuous(breaks = seq(rnk_thres[1], rnk_thres[2], by = 1)) +
  labs(x = "Rank", y = "Wealth (Billion)", title = "Wealth comparison")
ggplotly(rnk_bar)
```
```{r age-hist}
histo = region_select |>
  ggplot(aes(x = age, fill = region_code)) +
  geom_histogram(binwidth = 2) +
  facet_grid(region_code ~ .) + 
  theme(legend.position = "none")
ggplotly(histo)
```


#### Pie

```{r gender-pie, include=F}
region_select |>
  group_by(gender, region_code) |>
  summarise(count = n()) |>
  ggplot(aes(x = "", y = count, fill = gender)) +
  geom_bar(stat = "identity", width = 1, color = "white") +
  coord_polar("y", start = 0) +
  facet_grid(. ~ region_code) +
  scale_fill_manual(values = c("tomato", "lightblue", "grey")) +
  theme_void() +
  theme(legend.position = "right") 
```

```{r industries-pie, include = F}
region_select |>
  group_by(industries, region_code) |>
  summarise(count = n()) |>
  ggplot(aes(x = "", y = count, fill = industries)) +
  geom_bar(stat = "identity", width = 1, color = "white") +
  coord_polar("y", start = 0) +
  facet_grid(. ~ region_code) +
  theme_void() +
  theme(legend.position = "right") +
  viridis::scale_fill_viridis(
    name = "industries", 
    discrete = TRUE,
    option = "H"
  )
```

```{r self-made-pie, include = F}
region_select |>
  group_by(self_made, region_code) |>
  summarise(count = n()) |>
  ggplot(aes(x = "", y = count, fill = self_made)) +
  geom_bar(stat = "identity", width = 1, color = "white", alpha = .8) +
  coord_polar("y", start = 0) +
  facet_grid(. ~ region_code) +
  scale_fill_manual(values = c("orange", "violet", "grey")) +
  theme_void() +
  theme(legend.position = "right")

```


## Deployment
These geographic analysis is visualized via Shiny App and deployed to [Shiny App](https://tptrix29.shinyapps.io/p8105-geo-shiny/). 


