---
title: "Hypothesis Testing"
author: "Sitian Zhou, Mengxiao Luan, Shuchen Dong"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: true
    toc_float: true
    code_folding: "hide"
---

Comments:

1. ANOVA assumptions violated (?) even after transformation b/c the test is too sensitive with large sample size

2. need to combine with results in longitudinal EDA to determine which year/industry to look at

```{r setup, include=FALSE}
library(tidyverse)
library(multcompView)
library(EnvStats)

knitr::opts_chunk$set(
  echo = TRUE,
  warning = FALSE,
  message = FALSE,
  comment = NA
  )

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d

theme_set(theme_minimal() + theme(legend.position = "bottom"))
```


```{r import & subset data}
bil_data <-
  read_csv("../data/billionaire_gdp.csv")

bil_13_23 <-
  bil_data |> 
  filter(year == 2013 | year == 2023)
```

## Two sample prop test: Compare the proportion of female billionaires in 2013 and 2023

$H_0: p_{female\_2013} = p_{female\_2023}$ vs. $H_1: p_{female\_2013} \neq p_{female\_2023}$


```{r female billionaires summary}
bil_gender_sum <-
  bil_13_23 |>
  drop_na(gender) |> 
  group_by(year) |> 
  summarize(
    total = n(),
    female = sum(gender == "Female"))
bil_gender_sum|> 
  knitr::kable()
```


```{r female billionaires prop.test}
# perform 2-sample prop test for proportion of female billionaires in year 2013 and 2023
prop.test(c(filter(bil_gender_sum, year == 2013) |> 
              pull(female), filter(bil_gender_sum, year == 2023) |> pull(female)), 
          n = c(filter(bil_gender_sum, year == 2013) |> 
                  pull(total), filter(bil_gender_sum, year == 2023) |> pull(total))) |> 
  broom::tidy() |> 
  knitr::kable()
```


## Two sample prop test: Compare proportions of billionaires in health care industry in 2013 and 2023

$H_0: p_{healthcare\_2013} = p_{healthcare\_2023}$ vs. $H_1: p_{healthcare\_2013} \neq p_{healthcare\_2023}$

```{r industry summary}
bil_indus_sum <-
  bil_13_23 |> 
  group_by(year) |> 
  summarize(
    total = n(),
    indus = sum(industries == "Health care"))
bil_indus_sum|> 
  knitr::kable()
```


```{r prop.test for the industry in 2013 and 2023}
prop.test(x = c(filter(bil_indus_sum, year == 2013) |> 
              pull(indus), filter(bil_indus_sum, year == 2023) |> pull(indus)), 
          n = c(filter(bil_indus_sum, year == 2013) |> 
              pull(total), filter(bil_indus_sum, year == 2023) |> pull(total))) |> 
  broom::tidy() |> 
  knitr::kable()
```

## ANOVA: compare the net worth among multiple age groups in 2023

### Diagnostics

```{r anova diagnostics}
bil_age <-
  bil_data |> 
  drop_na(age) |> 
  filter(year == 2023) |> 
  mutate(
    age_group = 
      case_when(
        age <= 55 ~ "55 and below",
        age > 55 & age <= 65 ~ "55~65",
        age > 65 & age <= 75 ~ "65~75",
        age > 75 ~ "over 75"),
    age_group = 
      fct_relevel(age_group, c("55 and below", "55~65", "65~75", "over 75")))
# distribution of residuals
model_ori = lm(net_worth ~ age_group, data = bil_age)

model_ori[[2]] |> 
  as.data.frame() |> 
  rename("residuals" = "model_ori[[2]]") |> 
  ggplot(aes(x = residuals)) + geom_histogram() +
  labs(x = "residual", title = "Histogram of errors")

# boxplot
bil_age |> 
  ggplot(aes(x = age_group, y = net_worth, fill = age_group)) + geom_boxplot()
```

From the boxplot above, we can tell that the distribution of `net_worth` in each age group is right-skewed, which violates the normality and homoscedasticity assumptions of ANOVA. Hence, we have to transform the data before performing the ANOVA test.

### Data Transformation


```{r transformation}
# use boxcox method to transform the data
L1 = boxcox(model_ori ,objective.name = "Shapiro-Wilk",optimize = TRUE)$lambda

bil_age <-
  bil_age |> 
  mutate(
    net_worth_tran = (net_worth^(L1)-1)/L1)

```

### Diagnostics After Transformation

```{r anova diagnostics post-tran}
# distribution of residuals
model_t = lm(net_worth_tran ~ age_group, data = bil_age)

model_t[[2]] |> 
  as.data.frame() |> 
  rename("residuals" = "model_t[[2]]") |> 
  ggplot(aes(x = residuals)) + geom_histogram() +
  labs(x = "residual", title = "Histogram of errors")

# boxplot
bil_age |> 
  ggplot(aes(x = age_group, y = net_worth_tran, fill = age_group)) + geom_boxplot()

```

### ANOVA Results

$H_0: \mu_{55\space and \space below} = \mu_{55-65} = \mu_{65-75} = \mu_{over \space 75}$ vs.
$H_1:$ at least one $\mu$ differs

```{r anova results}
summary(model_t) |> broom::glance()
```

Since the test result of ANOVA is significant, we want to know which group(s) have different mean `net_worth_tran` from the rest.

### Pairwise Comparisons

```{r pairwise comparisons}
res <- aov(net_worth_tran ~ age_group, data = bil_age)
TukeyHSD(res) |> broom::tidy()

```

#### Visualize the result in a boxplot
```{r results visualization}
letter <- multcompLetters4(res, TukeyHSD(res))
letter_df <- as.data.frame.list(letter[[1]])
dt <- 
  bil_age |> 
  group_by(age_group) |> 
  summarize(pos = quantile(net_worth_tran)[4]) |>  
  mutate(letter = letter_df |> pull(Letters))

bil_age |> 
  ggplot(aes(x = age_group, y = net_worth_tran, fill = age_group)) + 
  geom_boxplot() +
  geom_text(data = dt, aes(x = age_group, y = pos, label = letter), 
            hjust = -1.25, vjust = -0.8)
```





