---
title: "EDA_Longitudinal"
author: "Shuchen Dong, Sitian Zhou, Mengxiao Luan"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: true
    toc_float: true
    code_folding: "hide"
---

```{r lib-import, message = FALSE}
library(tidyverse)
library(readxl)
library(dplyr)
library(ggplot2)
library(plotly)
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(
	echo = TRUE,
	warning = FALSE,
	message = FALSE,
	fig.width = 8, 
  fig.height = 6,
  out.width = "90%"
)

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d

theme_set(
  theme_bw() +
  theme(plot.title = element_text(hjust = 0.5),
        text = element_text(family = "Helvetica")) + 
  theme(legend.position = "right"))
```

```{r data}
# import data
billionaire_gdp_indus_usa <- read.csv("../data/billionaire_gdp_indus_usa.csv")
billionaire_gdp <- read.csv("../data/billionaire_gdp.csv")
gdp <- read.csv("../data/gdp.csv")
```



## Economic trends
It explores the dynamics of economies, industries, and cities in terms of GDP and wealth concentration.

### Country economy
GDP Change from 2010 to 2022 (Countries: USA, CHN, RUS, JPN, DEU):
```{r 5 country gdp}
countries <- c("USA", "CHN", "RUS", "JPN", "DEU")

gdp_selected <- 
  gdp[gdp$code %in% countries & gdp$year >= 2010 & gdp$year <= 2022, ]

# draw GDP plot
gdp_con =
  ggplot(data = gdp_selected, aes(x = year, y = gdp, color = code)) +
  theme_minimal() +
  geom_line(size = 1.2) +
  labs(title = "GDP change from 2010 to 2022", x = "Year", y = "GDP", color = "Country Code") +
  scale_x_continuous(breaks = 2010:2022)+
  theme(plot.title = element_text(hjust = 0.5)) 

ggplotly(gdp_con)

```

Summary: The graph demonstrates the varying economic trajectories of these five countries over a 12-year period, with China showing the most remarkable growth rate, the USA maintaining a strong and steadily increasing GDP, while Japan shows relatively little change, indicating different stages of economic development and growth rates.


### Industry

Number of billionaires in different industries by year (2010 to 2023):
```{r barplot-all industries}
bil_industry = 
  billionaire_gdp |> 
  mutate(industries = replace(industries, industries == "Billionaire", NA))|>
  drop_na(industries) |> 
  group_by(year, industries) |> 
  summarize(n = n()) |> 
  ggplot(aes(x = year, y = n, fill = industries))+
  theme_minimal()+
  geom_bar(position="stack", stat="identity")+
  labs(title = "GDP change from 2010 to 2023", x = "Year", y = "GDP", color = "Country Code")+
  scale_x_continuous(breaks = 2010:2023)+
  theme(plot.title = element_text(hjust = 0.5)) 

ggplotly(bil_industry)
```


Number of billionaires in different industries by year in the USA (2017 to 2022):
```{r barplot-US industries}
indus_usa = 
  billionaire_gdp_indus_usa |> 
  drop_na(industry_gdp) |> 
  group_by(year, industries) |> 
  summarize(n = n()) |> 
  ggplot(aes(x = year, y = n, fill = industries)) +
  theme_minimal() +
  geom_bar(position="stack", stat="identity") +
  labs(title = "GDP change in the USA from 2017 to 2022", x = "Year", y = "GDP", color = "Country Code") +
  scale_x_continuous(breaks = 2017:2022)+
  theme(plot.title = element_text(hjust = 0.5)) 

ggplotly(indus_usa)

```


Total industry GDP by year in the USA (2017 to 2022):
```{r US industry gdp overtime}
data_filtered <- billionaire_gdp_indus_usa %>%
  filter(year >= 2017 & year <= 2022) %>% 
  na.omit()


total_gdp_by_year_industry <- data_filtered %>%
  group_by(year, industries) %>%
  summarise(total_gdp_by_year_industry = sum(industry_gdp, na.rm = TRUE))

gdp_industry_plot <- total_gdp_by_year_industry |> 
  ggplot(aes(x = year, y = total_gdp_by_year_industry, group = industries, color = industries)) +
    geom_line(size = 1.2) +
    theme_minimal() +
    labs(title = "Total industry GDP by year from 2017 to 2022 in USA", x = "Year", y = "Industry GDP", color = "Industries")+
  theme(plot.title = element_text(hjust = 0.5))

ggplotly(gdp_industry_plot)

```
 
Summary: It shows that from 2017 to 2022, the Construction & Engineering and Finance & Investment sectors in the USA experienced the most significant growth. In contrast, the Media and Entertainment industry saw the least growth. It has been observed that due to the impact of the COVID-19 pandemic, billionaires in all industries in the USA and around the world faced an overall slump in 2020. Despite such a setback, these indicators generally tend to rise over time, indicating that the GDP of each industry is growing comprehensively.


### City
Top 5 Cities' Net Worth Trend by Year (2010 to 2023):
find different city of residence billionaire net worth change over time
```{r net_worth vs cities overtime}
data_filtered <- billionaire_gdp %>%
  filter(
    year >= 2010 & year <= 2023, 
    country_of_citizenship == "United States")%>%
  na.omit()

net_worth_by_city_year <- data_filtered %>%
  group_by(year, city_of_residence) %>%
  summarise(total_net_worth = sum(net_worth, na.rm = TRUE)) %>%
  ungroup()

# top5
top_cities_by_year <- net_worth_by_city_year %>%
  group_by(year) %>%
  slice_max(order_by = total_net_worth, n = 5) %>%
  ungroup()

tnw_residence = 
  ggplot(top_cities_by_year, aes(x = year, y = total_net_worth, group = city_of_residence, color = city_of_residence)) +
  geom_line(size = 1.2) +
  theme_minimal() +
  labs(title = "Top 5 cities' net worth trend by year (2010-2023)",
       x = "Year", y = "Total Net Worth (Billions)",
       color = "City of Residence") +
  theme(legend.position = "right") +
  theme(plot.title = element_text(hjust = 0.5))

ggplotly(tnw_residence)

```

Note: Here are the top 5, and there are indeed 5 more. This is because in some years, the total city of net worth is surpassed by others.

Summary: The chart shows rising billionaire net worth in five cities from 2010 to 2023, with New York's significant growth standing out. Dallas and Medina also grew notably post-2015, while Austin and Bentonville had modest increases. A slight dip in 2020 suggests the impact of the COVID-19 pandemic. There's a data gap for Bentonville in 2020 and a title discrepancy with ten cities listed instead of five, indicating wealth concentration and potential urban-rural disparities.



## Wealth Distribution
This category focuses on the distribution of wealth among different groups, genders, and the rise of self-made billionaires.

### Global

A quick overview: trend of total net worth of all billionaires (2010 to 2023)
```{r networth all overtime}
bil_all = 
  billionaire_gdp |> 
  group_by(year) |> 
  summarize(total_net_worth = sum(net_worth)) |> 
  ggplot(aes(x = year, y = total_net_worth)) +
  geom_line(size = 1.2) +
  theme_minimal() +
  labs(title = "Total net worth of billionaire from 2010 to 2023",
       x = "Year", y = "Total Net Worth (Billions)",
       color = "Country of Citizenship")+
  geom_point() +
  scale_x_continuous(breaks = 2010:2023)+
  theme(plot.title = element_text(hjust = 0.5))

ggplotly(bil_all)

```

Summary: Total net worth of all billionaires plummeted in 2016 and 2020, and then skyrocketed in 2021.

### Country

Total Net Worth of Billionaires (2010 to 2023, Top 5 Countries):
```{r networth 5 countries overtime}
data_filtered <- billionaire_gdp %>%
  filter(year >= 2010 & year <= 2023)

# group by country net worth over time 
total_net_worth_by_country <- data_filtered %>%
  group_by(country_of_citizenship) %>%
  summarise(total_net_worth = sum(net_worth, na.rm = TRUE))

# net worth top 5 countries
top_countries <- total_net_worth_by_country %>%
  top_n(5, total_net_worth) %>%
  pull(country_of_citizenship)

top_countries_data <- data_filtered %>%
  filter(country_of_citizenship %in% top_countries)

total_net_worth_by_year_country <- top_countries_data %>%
  group_by(year, country_of_citizenship) %>%
  summarise(total_net_worth = sum(net_worth, na.rm = TRUE))

# plot
tnw_citizen =
  ggplot(total_net_worth_by_year_country, aes(x = year, y = total_net_worth, group = country_of_citizenship, color = country_of_citizenship)) +
  geom_line(size = 1.2) +
  theme_minimal() +
  labs(title = "Total net worth of billionaire by top 5 countries from 2010 to 2023",
       x = "Year", y = "Total Net Worth (Billions)",
       color = "Country of Citizenship")+
  theme(plot.title = element_text(hjust = 0.5))

ggplotly(tnw_citizen)
```

Summary: Shows the combined net worth of billionaires in the top 5 countries.
Insight: Provides a perspective on how billionaire wealth is distributed among leading economic nations, possibly reflecting on the economic health and wealth creation capabilities of these countries.

### Gender

Percentage of Billionaires by Gender (2010-2022):
```{r barplot-gender-scaled}
# scale it to 1 to analyze the proportions
bil_gender_prop <-
  billionaire_gdp|> 
  drop_na(gender) |> 
  group_by(year, gender) |> 
  summarize(n = n()) |> 
  group_by(year) |> 
  mutate(total = sum(n),
         prop = n/ total)

bil_gender = 
  bil_gender_prop |> 
  ggplot(aes(title = "Proportion of billionaires by gender from 2010 to 2022", x = year, y = prop, fill = gender))+
  geom_bar(position="stack", stat="identity") +
  scale_x_continuous(breaks = 2010:2023)

ggplotly(bil_gender)
  
```
Comments: The first chart, a bar plot, illustrates the proportion of male and female billionaires from 2010 to 2023, indicating a consistent predominance of males in each year, with females representing a very small fraction of the total. This disparity in representation may reflect long-standing gender imbalances in wealth accumulation and opportunities for economic advancement.

Total Net Worth by Gender (2010 to 2022):
```{r ave_net_worth by gender}

billionaires_filtered <- billionaire_gdp %>%
  filter(year >= 2010 & year <= 2022) %>%  na.omit()

total_net_worth_by_year_gender <- billionaires_filtered %>%
  group_by(year, gender) %>%
  summarise(total_net_worth = sum(net_worth, na.rm = TRUE))


tnw_gender = 
  ggplot(total_net_worth_by_year_gender, aes(x = year, y = total_net_worth, color = gender)) +
  geom_line(size = 1.2) +
  theme_minimal() +
  labs(title = "Total net worth by gender from 2017 to 2022",
       x = "Year", y = "Total Net Worth (in billions)",
       color = "Gender")+
  theme(plot.title = element_text(hjust = 0.5))

ggplotly(tnw_gender)
```
Comments: The second chart, a line graph, shows the total net worth by gender from 2010 to 2022. The net worth of male billionaires is significantly higher than that of females throughout the period, with a particularly steep increase for males after 2015. The consistent gap and the sharp rise in male net worth could be due to a combination of factors, including the compounding effect of existing wealth, gender disparities in key industries, and potentially more males being active in high-growth sectors such as technology and finance.

Summary: Both charts underscore a persistent gender gap in billionaire wealth, suggesting systemic barriers to entry and growth for women in the highest echelons of wealth. The trends could also reflect broader societal and economic structures that historically favor male dominance in business and wealth creation.

### Self-Made Status Count Over Years:
```{r barplot-self_made-scaled}
# scale it to 1 to analyze the proportions
bil_selfmade_prop <-
  billionaire_gdp|> 
  drop_na(self_made) |> 
  group_by(year, self_made) |> 
  summarize(n = n()) |> 
  group_by(year) |> 
  mutate(total = sum(n),
         prop = n/ total)

bil_selfmade = 
  bil_selfmade_prop |> 
  ggplot(aes(x = year, y = prop, fill = self_made)) +
  geom_bar(position="stack", stat="identity") +
  scale_x_continuous(breaks = 2010:2023)

ggplotly(bil_selfmade)
```

Summary: Counts the number of self-made billionaires over time.
Insight: Reflects the trends in entrepreneurial success and self-made wealth, which could be indicative of the broader economic opportunities and the state of the entrepreneurial ecosystem.
Each of these plots provides valuable insights into specific aspects of economic trends and wealth distribution. Together, they paint a comprehensive picture of how wealth is created, distributed, and evolves over time across different countries, industries, genders, and cities.

## Summary

GDP-Temporal Wealth Patterns:
Examine the rise and fall of billionaire wealth across different economic cycles, including booms, busts, and recoveries.







