---
title: "Longitudinal EDA"
author: "Shuchen Dong, Sitian Zhou, Mengxiao Luan"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: true
    toc_float: true
    code_folding: "hide"
---

```{r lib-import, include=FALSE}
library(tidyverse)
library(readxl)
library(dplyr)
library(ggplot2)
library(plotly)
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(
	echo = TRUE,
	warning = FALSE,
	message = FALSE,
	fig.width = 8, 
  fig.height = 6,
  out.width = "90%"
)

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d

theme_set(
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5),
        text = element_text(family = "Helvetica")) + 
  theme(legend.position = "right"))
```

```{r data}
# import data
billionaire_gdp_indus_usa <- read.csv("../data/billionaire_gdp_indus_usa.csv")
billionaire_gdp <- read.csv("../data/billionaire_gdp.csv")
gdp <- read.csv("../data/gdp.csv")
```

**Global Billionaires' Wealth Trends Analysis**

We conduct some exploratory analysis from two main aspects:

* Billionaire Economic Trends:

  - *GDP Trends:* Analysis of changes in GDPs of 5 countries home to global billionaires, highlighting growth rates and economic development stages.

  - *Sectoral Wealth Growth:* Examination of the wealth growth within industries dominated by billionaires, noting the rise of new economic powerhouses.

  - *Millionaire Residency Wealth Analysis:* Tracking the net worth trends in Millionaire Residency with the highest concentration of billionaires, identifying urban patterns of wealth accumulation.


* Billionaire Wealth Distribution  

  - *Global Net Worth Trends:* Observation of the global net worth changes of billionaires, highlighting significant fluctuations due to global economic conditions.

  - *Region Wealth Comparison:* Analysis of the combined net worth of billionaires in the top five regions, reflecting the economic health and wealth generation in these nations.

  - *Gender Wealth Gap:* Examination of the distribution and total net worth of billionaires by gender, underscoring the persistent gender gap in billionaire wealth.
  
  - *Self-Made Billionaires:* Tracking the rise of self-made billionaires over the years, indicating the evolving landscape of entrepreneurial success and self-made wealth.



## Billionaire Economic trends
It explores the dynamics of economies, industries, and residences in terms of GDP and wealth concentration.

### GDP Trends
```{r 5 country gdp overtime}
countries <- c("USA", "CHN", "RUS", "JPN", "DEU")

gdp_selected <- 
  gdp[gdp$code %in% countries & gdp$year >= 2010 & gdp$year <= 2022, ]

# draw GDP plot
gdp_con =
  ggplot(data = gdp_selected, aes(x = year, y = gdp, color = code)) +
  theme_minimal() +
  geom_line(size = 1.2) +
  labs(title = "GDP change from 2010 to 2022", x = "Year", y = "GDP(trillions)", color = "Region Code") +
  scale_x_continuous(breaks = 2010:2022)+
  theme(plot.title = element_text(hjust = 0.5),
        text = element_text(family = "Helvetica")) 

ggplotly(gdp_con)

```

**The graph demonstrates the varying economic trajectories of these five countries(USA, CHN, RUS, JPN, DEU) over a 12-year period, with China showing the most remarkable growth rate, the USA maintaining a strong and steadily increasing GDP, while Japan shows relatively little change.**


### Sectoral Wealth Growth
```{r barplot-all industries overtime}
bil_industry = 
  billionaire_gdp |> 
  mutate(industries = replace(industries, industries == "Billionaire", NA))|>
  drop_na(industries) |> 
  group_by(year, industries) |> 
  summarize(n = n()) |> 
  ggplot(aes(x = year, y = n, fill = industries))+
  theme_minimal()+
  geom_bar(position="stack", stat="identity")+
  labs(title = "Number of billionaires in different industries from 2010 to 2023", 
       x = "Year", y = "Number of billionaires", fill = "Industries")+
  scale_x_continuous(breaks = 2010:2023)+
  theme(plot.title = element_text(hjust = 0.6),
        text = element_text(family = "Helvetica")) 

ggplotly(bil_industry)
```

**The global chart shows a clear upward trend in the number of billionaires across various industries, with significant growth in sectors such as Technology and Finance & Investments.** 

```{r barplot-US industries overtime}
indus_gdp = 
  billionaire_gdp_indus_usa |> 
  drop_na(industry_gdp) |> 
  group_by(year, industries) |> 
  summarize(n = n()) |> 
  ggplot(aes(x = year, y = n, fill = industries)) +
  theme_minimal() +
  geom_bar(position="stack", stat="identity") +
  labs(title = "Number of billionaires in different industries in the USA from 2017 to 2022", 
       x = "Year", y = "Number of billionaires", fill = "Industries") +
  scale_x_continuous(breaks = 2017:2022)+
  theme(plot.title = element_text(hjust = 0.6),
        text = element_text(family = "Helvetica")) 

ggplotly(indus_gdp)

```

**The USA-specific chart also displays an increase in the number of billionaires, particularly in Technology and Finance & Investments, while those engaged in health care has increased significantly after 2020 due to the global epidemic.**


```{r US industry gdp overtime}
data_filtered <- billionaire_gdp_indus_usa %>%
  filter(year >= 2017 & year <= 2022) %>% 
  na.omit()

gdp_industry_usa <- data_filtered |> 
  select(year, industries, industry_gdp) |> 
  unique() |> 
  ggplot(aes(x = year, y = industry_gdp, color = industries)) +
  geom_line(size = 1.2) +
  theme_minimal() +
  labs(title = "Industry GDP by year from 2017 to 2022 in the USA", 
         x = "Year", y = "Industry GDP(trillions)", color = "Industries")+
  theme(plot.title = element_text(hjust = 0.5),
        text = element_text(family = "Helvetica"))

ggplotly(gdp_industry_usa)
```

**The chart illustrates the GDP changes of various industries in the USA from 2017 to 2022. It highlights that the Real Estate sector experienced the most significant growth, while the Media and Entertainment industry saw the least. There was a noticeable downturn in 2020 attributed to the COVID-19 pandemic, affecting billionaires across all sectors. **


### Millionaire Residency Wealth Analysis
```{r}
data_filtered <- billionaire_gdp |> 
  filter(year >= 2010 & year <= 2023, 
    country_of_citizenship == "United States")

net_worth_by_city_year <- data_filtered |> 
  select(year, city_of_residence, net_worth) |> 
  unique() |> 
  drop_na(city_of_residence)

# top5 total_net_worth
top_cities_by_year <- net_worth_by_city_year %>%
  group_by(year) %>%
  slice_max(order_by = net_worth, n = 5) %>%
  ungroup()

nw_residence = 
  ggplot(top_cities_by_year, aes(x = year, y = net_worth, color = city_of_residence)) +
  geom_line(size = 1.2) +
  labs(title = "Net Worth Evolution in Top 5 Cities by Millionaire Residency (2010-2023)",
       x = "Year", y = "Net worth(billions)",
       color = "City of residence") +
  scale_x_continuous(breaks = 2010:2023)+
  theme(legend.position = "right") +
  theme(plot.title = element_text(hjust = 0.5),
        text = element_text(family = "Helvetica"))

ggplotly(nw_residence)

```

*Note: In this analysis, 11 cities are considered, with annual changes in rankings due to variations in net worth. Net worth serves as the 5 sorting criterion and the decline noted in 2023 is likely a result of incomplete data.*

**The data indicates that while cities like Medina and Omaha remain consistent high performers, there's a dynamic shift with other cities like Austin and Seattle showing rapid growth, indicating a possible diversification in the geographical distribution of wealth.**


## Billionaire Wealth Distribution
This category focuses on the distribution of wealth among different groups, genders, and the rise of self-made billionaires.

### Global Net Worth Trends
```{r networth all overtime}
bil_all = 
  billionaire_gdp |> 
  group_by(year) |> 
  summarize(total_net_worth = sum(net_worth)) |> 
  ggplot(aes(x = year, y = total_net_worth, color = "#440154")) +
  geom_line(size = 1.2) +
  labs(title = "Total net worth of billionaire from 2010 to 2023",
       x = "Year", y = "Total net worth (billions)")+
  scale_x_continuous(breaks = 2010:2023)+
  theme(plot.title = element_text(hjust = 0.5),
        legend.position = "none",
        text = element_text(family = "Helvetica"))

ggplotly(bil_all)

```

**The line graph portrays the trajectory of a billionaire's net worth from 2010 to 2023, with a general upward trend punctuated by notable declines in 2016 and 2020, likely due to market downturns and the economic impact of the COVID-19 pandemic, respectively. Following the 2020 dip, there was a dramatic rebound, culminating in a peak in 2021.**


### Region Wealth Comparison
```{r networth 5 regions overtime}
data_filtered <- billionaire_gdp %>%
  filter(year >= 2010 & year <= 2023)

# group by region_code net worth over time 
total_net_worth_by_region <- data_filtered %>%
  group_by(region_code) %>%
  summarise(total_net_worth = sum(net_worth, na.rm = TRUE))

# net worth top 5 regions
top_regions <- total_net_worth_by_region %>%
  top_n(5, total_net_worth) %>%
  pull(region_code)

top_regions_data <- data_filtered %>%
  filter(region_code %in% top_regions)

total_net_worth_by_year_region <- top_regions_data %>%
  group_by(year, region_code) %>%
  summarise(total_net_worth = sum(net_worth, na.rm = TRUE))

# plot
tnw_citizen =
  ggplot(total_net_worth_by_year_region, aes(x = year, y = total_net_worth, color = region_code)) +
  geom_line(size = 1.2) +
  theme_minimal() +
  labs(title = "Total net worth of billionaires by top 5 regions from 2010 to 2023",
       x = "Year", y = "Total net worth (billions)",
       color = "Region code")+
  scale_x_continuous(breaks = 2010:2023)+
  theme(plot.title = element_text(hjust = 0.5),
        text = element_text(family = "Helvetica"))

ggplotly(tnw_citizen)
```

**The graph comparing the net worth of billionaires from 2010 to 2023 in the top five regions shows the USA leading with a significant surge post-2015. China's trend suggests rapid wealth creation with some instability. Germany displays steady progress. India's consistent rise aligns with its economic development, while Hong Kong's initial growth is followed by a decline post-2019. This distribution of billionaire wealth highlights the varying economic dynamics and the impact of regional and global factors on wealth accumulation.**

### Gender Wealth Gap
```{r barplot-gender-scaled overtime}
# scale it to 1 to analyze the proportions
bil_gender_prop <-
  billionaire_gdp|> 
  drop_na(gender) |> 
  group_by(year, gender) |> 
  summarize(n = n()) |> 
  group_by(year) |> 
  mutate(total = sum(n),
         prop = n/ total)

bil_gender = 
  bil_gender_prop |> 
  ggplot(aes(x = year, y = prop, fill = gender))+
  geom_bar(position="stack", stat="identity") +
  labs(title = "Proportion of billionaires by gender from 2010 to 2023",
       x = "Year", y = "Proportion", fill = "Gender")+
  scale_x_continuous(breaks = 2010:2023)+
  theme(plot.title = element_text(hjust = 0.5),
        text = element_text(family = "Helvetica"))

ggplotly(bil_gender)
  
```
**The bar chart illustrates the proportion of male and female billionaires from 2010 to 2023, indicating a consistent predominance of males in each year, with females representing a very small fraction of the total. This disparity in representation may reflect long-standing gender imbalances in wealth accumulation and opportunities for economic advancement.**


```{r ave_net_worth by gender overtime}
billionaires_filtered <- billionaire_gdp %>%
  filter(year >= 2010 & year <= 2023) %>%  na.omit()

total_net_worth_by_year_gender <- billionaires_filtered %>%
  group_by(year, gender) %>%
  summarise(total_net_worth = sum(net_worth, na.rm = TRUE))


tnw_gender = 
  ggplot(total_net_worth_by_year_gender, aes(x = year, y = total_net_worth, color = gender)) +
  geom_line(size = 1.2) +
  theme_minimal() +
  labs(title = "Total net worth by gender from 2010 to 2023",
       x = "Year", y = "Total net worth (billions)", color = "Gender")+
  scale_x_continuous(breaks = 2010:2023)+
  theme(plot.title = element_text(hjust = 0.5),
        text = element_text(family = "Helvetica"))

ggplotly(tnw_gender)
```

**The second line graph shows the net worth by gender from 2010 to 2023. The net worth of male billionaires is significantly higher than that of females throughout the period, with a particularly steep increase for males after 2015. The consistent gap and the sharp rise in male net worth could be due to a combination of factors, including the compounding effect of existing wealth, gender disparities in key industries, and potentially more males being active in high-growth sectors such as technology and finance.**


### Self-Made Billionaires
```{r barplot-self_made-scaled overtime}
# scale it to 1 to analyze the proportions
bil_selfmade_prop <-
  billionaire_gdp|> 
  filter(year > 2010) |> 
  drop_na(self_made) |> 
  group_by(year, self_made) |> 
  summarize(n = n()) |> 
  group_by(year) |> 
  mutate(total = sum(n),
         prop = n/ total)

bil_selfmade = 
  bil_selfmade_prop |> 
  ggplot(aes(x = year, y = prop, fill = self_made)) +
  geom_bar(position="stack", stat="identity") +
  labs(title = "Proportion of self-made status count from 2011 to 2023",
       x = "Year", y = "Proportion", fill = "Self-made")+
  theme(plot.title = element_text(hjust = 0.5))+
  scale_x_continuous(breaks = 2011:2023)

ggplotly(bil_selfmade)
```

*Note: Remove data for the year 2010 due to errors*

**The bar chart depicting the proportion of self-made billionaires from 2011 to 2023 demonstrates a consistent majority of individuals who have amassed their wealth independently, without inheritance or existing family wealth. **







